---
title: "Data Set"
author: "Adrian Harris"
date: "12/16/2019"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r}
library(readr)
library(ggplot2)
library(wesanderson)
```

```{r}
anth_data_set_ <- read_csv("~/Desktop/anth data set .csv")
```

```{r}
ant = anth_data_set_ 
```

```{r}
colnames(ant) = c("Age", "Sex", "Location", "Sexual Partners",
                  "Level of Sexual Education", 
                  "Protection", "Race", "Ethnicity", "Prep Knowledge", 
                  "Sexual Education Improtance Level","Who Should Be Sexual an Educator", 
                  "Methods of Birth Control", 
                  "Preventive Methods of STDs/STI", "Final Comments", "Name", 
                  "Sexual Education Stigma", "Sexual Orientation")
```

```{r}
contin <- wes_palette("Darjeeling1", 100, type = "continuous")
```

### Graph one 

Ages 20 - 28. There are two from the same town and all within a 10-mile radius.

```{r}
ggplot(ant) +
  aes(x = ant$Location, fill = Age)+
  theme_bw()+
  geom_bar(position = "dodge")+
  scale_color_discrete(name = "Age")+
  scale_fill_gradientn(colours = contin)+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(title = "Obevastion vs Age",
       x = "Location",
       y = "Counts")

```

### Graph two 

Mixed field 
```{r}
ggplot(ant) +
  aes(x = ant$Location, fill = Ethnicity)+
  theme_bw()+
  geom_bar(position = "dodge")+
  scale_color_discrete(name = "Ethnicity")+
  xlab("Location")+
  ylab("Counts")+
  ggtitle("Obevastion vs Ethnicity")+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(values = wes_palette("Moonrise3", n = 5))
```


### Graph three 

Mostly Caucasian 
```{r}
ggplot(ant) +
  aes(x = ant$Location, fill = Race)+
  theme_bw()+
  geom_bar(position = "dodge")+
  xlab("Location")+
  ylab("Counts")+
  ggtitle("Obevastion vs Race")+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(values = wes_palette("Moonrise3", n = 5))
```


### Graph four 

Mixed group 
```{r}
ggplot(ant) +
  aes(x = ant$Location, fill = `Level of Sexual Education`)+
  theme_bw()+
  geom_bar(position = "dodge")+
  xlab("Location")+
  ylab("Counts")+
  ggtitle("Obevastion vs Level of Sexual Education")+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(values = wes_palette("Moonrise3", n = 5))
```

### Graph five

Half said yes. One with Highschool education on sexual education said he rarely uses protection. 

```{r}
ggplot(ant) +
  aes(x = ant$Location, fill = Protection)+
  theme_bw()+
  geom_bar(position = "dodge")+
  xlab("Location")+
  ylab("Counts")+
  ggtitle("Obevastion vs Protection Usage")+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(values = wes_palette("Moonrise3", n = 5))
```

### Graph six 

One with the master education knows about PRep+
```{r}
ggplot(ant) +
  aes(x = ant$Location, fill = `Prep Knowledge`)+
  theme_bw()+
  geom_bar(position = "dodge")+
  xlab("Location")+
  ylab("Counts")+
  ggtitle("Obevastion vs PRep+ Knowledge")+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(values = wes_palette("Moonrise3", n = 5))
```





### Graph seven 

Most know about birth control methods 




```{r}
ggplot(ant) +
  aes(x = ant$Location, fill = `Methods of Birth Control`)+
  theme_bw()+
  geom_bar(position = "dodge")+
  xlab("Location")+
  ylab("Counts")+
  ggtitle("Obevastion vs Birth Control Knowledge")+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(values = wes_palette("Moonrise3", n = 5))

```






### Graph eight 

Teachers and Parents have about an even split 
```{r}

ggplot(ant) +
  aes(x = ant$Location, fill = `Who Should Be Sexual an Educator`)+
  theme_bw()+
  geom_bar(position = "dodge")+
  xlab("Location")+
  ylab("Counts")+
  ggtitle("Obevastion vs Who Should Be Sexual an Educator")+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(values = wes_palette("Moonrise3", n = 5))

```


### Graph nine 

Most said yes 
```{r}
ggplot(ant) +
  aes(x = ant$Location, fill = `Sexual Education Stigma`)+
  theme_bw()+
  geom_bar(position = "dodge")+
  xlab("Location")+
  ylab("Counts")+
  ggtitle("Obevastion vs Sexual Education Stigma")+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(values = wes_palette("Moonrise3", n = 5))
```

### Graph ten 

The observation that rarely uses protection is also a homosexual this means he can be a contributor to the spread of STDs/STIs in this area.
```{r}
ggplot(ant) +
  aes(x = ant$Location, fill = `Sexual Orientation`)+
  theme_bw()+
  geom_bar(position = "dodge")+
  xlab("Location")+
  ylab("Counts")+
  ggtitle("Obevastion vs Sexual Orientation")+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(values = wes_palette("Moonrise3", n = 5))
```

### Learning Algorithm

Logistic regression

```{r}
ant_code_ <- read_csv("~/Desktop/ant code .csv")
```

```{r}
antcode = ant_code_ 
```

```{r}
colnames(antcode) = c("Age", "Sex", "Location", "Sexual Partners",
                  "Level of Sexual Education", 
                  "Protection", "Race", "Ethnicity", "Prep Knowledge", 
                  "Sexual Education Improtance Level","Who Should Be Sexual an Educator", 
                  "Methods of Birth Control", 
                  "Preventive Methods of STDs/STI", "Final Comments", "Name", 
                  "Sexual Education Stigma", "Sexual Orientation")
```

```{r}
newantcode = antcode[,-15]
```

Prediction on stigma 
```{r}
logreg5 = glm(newantcode$`Sexual Education Stigma`~., 
              data = newantcode, family = binomial)
```

```{r}
pro89 = predict(logreg5, newantcode,  type = "response")
```

```{r}
predict16 = ifelse(pro89>.5,1,0)
```

Confusion matrix. The actuals are getting predicted as No's.
```{r}
table(predict16, newantcode$`Prep Knowledge`)
```

```{r}
levels(newantcode$`Sexual Education Stigma`) = c(0, 1)
newantcode$`Sexual Education Stigma` = as.factor(newantcode$`Sexual Education Stigma`)
```

Changing into a data frame so it can be plotted 
```{r}
antpre = data.frame(predict16, prob89 = pro89, 
                      Age = newantcode$Age,
                      Sex = newantcode$Sex,
                      Location = newantcode$Location,
                      Sex_Part = newantcode$`Sexual Partners`,
                      Level_sex = newantcode$`Level of Sexual Education`,
                      Protect = newantcode$Protection,
                      Race = newantcode$Race,
                      Ethic = newantcode$Ethnicity,
                      Prep = newantcode$`Prep Knowledge`,
                      Sex_level = newantcode$`Sexual Education Improtance Level`,
                      Educator = newantcode$`Who Should Be Sexual an Educator`,
                      Methods = newantcode$`Methods of Birth Control`,
                      Std = newantcode$`Preventive Methods of STDs/STI`,
                      Final = newantcode$`Final Comments`,
                      Stigma = newantcode$`Sexual Education Stigma`,
                      orin = newantcode$`Sexual Orientation`)

```

```{r}
ggplot(antpre)+
  aes(x = prob89, 
      y = newantcode$`Sexual Education Stigma`, 
      color = newantcode$`Sexual Education Stigma`)+
  theme_bw()+
  xlab("Probability")+
  ylab("Zero for no One for yes")+
  scale_color_discrete(name = 'Yes or No')+
  ggtitle("Logistic Regression on Stigma")+
  theme(plot.title = element_text(hjust = 0.5))+
  geom_point()+
  scale_fill_manual(values = wes_palette("Moonrise3", n = 2))
```


 Not enough *data* to perform the learning algorithm


















