---
title: "STA 3155 Project - Group 10"
author: "Adrian Harris,Butrint Komoni,Jing Lu,Robert Peri, and Shimeng Wang"
date: "5/13/2020"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

# Libraries 

```{r, results='hide'}

library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(wesanderson)
library(GGally)
library(gridExtra)
```

# Data Set

Hourly Interstate 94 Westbound traffic volume for MN DoT ATR station 301, roughly midway between Minneapolis and St Paul, MN. Hourly weather features and holidays included for impacts on traffic volume 

```{r}
metro = read_csv("~/Desktop/Metro_Interstate_Traffic_Volume (3).csv")
```

# Variables 

Holiday Categorical US National holidays plus regional holiday,Minnesota State Fair


Temp Numeric Average temp in kelvin


Rain_1h Numeric Amount in mm of rain that occurred in the hour



Snow_1h Numeric Amount in mm of snow that occurred in the hour


Clouds_all Numeric Percentage of cloud cover


Weather_main Categorical Short textual description of the current weather


Weather_description Categorical Longer textual description of the current weather


Date_time DateTime Hour of the data collected in local CST time - spilted 


Rraffic_volume Numeric Hourly I-94 ATR 301 reported westbound traffic volume

```{r}

```


# Pre-Processing 
```{r}
metro = separate(metro,col = date_time,into=c("Year", "Month", "Date"), sep = "-")
metro = separate(metro,col = Date,into=c("Date", "Time"), sep = " ")
metro = separate(metro,col = Time,into=c("Hour", "Minutes", "Seconds"), sep = ":")

metro$Month = as.numeric(metro$Month)
metro$Year = as.numeric(metro$Year)
metro$Date = as.numeric(metro$Date)
metro$Hour = as.numeric(metro$Hour)
metro$Minutes = as.numeric(metro$Minutes)
metro$Seconds = as.numeric(metro$Seconds)

metro$holiday = as.factor(metro$holiday)
metro$weather_main = as.factor(metro$weather_main)
metro$weather_description = as.factor(metro$weather_description)



```

# Picking our variables to use
Subsetting the dataset for fasting processing 

```{r}
metro2 = metro[1:1000,]
```

# Correlation Plot 
```{r}
ggpairs(metro2[,-7])
```

# Correlation Heat Map 
```{r}

ggcorr(metro2, label = T)
  
```

# Spilting the Data Set 

Train/Test/Validation Samples 
```{r}
set.seed(300)
sample = sample(3, nrow(metro), replace = T, prob = c(.6,.2,.2))
train = metro[sample ==1,]
test = metro[sample ==2,]
val = metro[sample ==3,]

```


## Simple Linear Regression - Part One 

# 1. Predicting traffic_volume based on hour

We picked traffic volume and hour because they have a relationship and they are somewhat on a normal distribution 

Question we ask was...Does nights and early mornings have less traffic volume and the opposite on during the day?
```{r}

```

# 2. Visualization

Weak correlation between variables, there is a relationship, but it's not linear. 

```{r}

ggplot(metro) + aes(metro$Hour, metro$traffic_volume, color = metro$weather_main)+
  geom_point()+
  theme_bw()+
  xlab("Hour")+
  ylab("Traffic Volume")+
  ggtitle("Hour vs Traffic Volume")+
  scale_color_discrete(name = 'Weather')+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(values = wes_palette("Moonrise3"))

```

# 3. Transformations

Hour Variable 
```{r}
mean(metro$Hour)
```

```{r}
ggplot(metro)+ aes(metro$Hour)+
  geom_histogram()+ 
  labs(title = "Hour by Count", 
       x = "Hour", y = "Count", 
       caption = "Dashed line represents the mean")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(values = wes_palette("Moonrise3"))+
  geom_vline(aes(xintercept=mean(metro$Hour)),
            color="blue", linetype="dashed", size=1)
```

Traffic Volume Variable  
```{r}
mean(metro$traffic_volume)
```


```{r}
ggplot(metro)+ aes(metro$traffic_volume)+ 
  geom_histogram()+
  labs(title = "Traffic Volume by Count", 
       x = "Traffic Volume", y = "Count", 
       caption = "Dashed line represents the mean")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))+
  geom_vline(aes(xintercept=mean(metro$Hour)),
            color="blue", linetype="dashed", size=1)
```


We performed multiple transformation to the volume data to move the numbers closers to the mean on a somewhat normal distribution. Looking at the individual variable no transformation was necessary. But in relation to the hour variable we need to transform the response. 

Since there is a upward curve and then downwards curve around hour 15 in relation to the volume  we applied a square root to account for the upward curve up to hour 15 and then raised traffic volume from  hour 16 to 24 to the second power.

metro$transform = ifelse(metro$Hour < 16, sqrt(metro$traffic_volume), 0)



metro$transform = ifelse(metro$Hour > 15, (metro$traffic_volume)^2, metro$transform)

This didn't improve the model in further analysis so we just went with taking the sqaure root of the repsonse variable. 

```{r}
ggplot(metro) + aes(metro$Hour, sqrt(metro$traffic_volume), color = metro$weather_main)+
  geom_point()+
  xlab("Hour")+
  ylab("Traffic Volume")+
  ggtitle("Hour vs Traffic Volume")+
  scale_color_discrete(name = 'Weather')+
  theme(plot.title = element_text(hjust = 0.5))
```


# 4. Model 


Predicted value of traffic volume = 39.00528(intercept) + 1.26529(hour)

Response is squared traffic volume 

Hour (b1) is the slope of the fitted regression line. Traffic volume is the predicted increase at 1.27 when the hour increases by one unit 

```{r}
fittedmodelone  = lm(sqrt(train$traffic_volume) ~ train$Hour, data = train)
```


# Plot one 

```{r}
plot(sqrt(train$traffic_volume)~train$Hour, main="Traffic in relation to Hour with Regression line",xlab="Hour",
        ylab="Sqrt of traffic Volume", )
abline(fittedmodelone,col="red")
```



```{r}

```

# 5.

Volume w/ hour predictor 

18.23 se which is high 

The R2 is 18.83 percent 

18.83% at the variation in the traffic volume can be explained in the linear association between traffic volume and the hour of the day. 



```{r}
summary(fittedmodelone)
```

# 6. 

Confidence interval for slope

95 Confidence that the true value of b1 is between 1.24 and 1.30 
```{r}
confint(fittedmodelone,level=0.95)
```

# 7. 

## Test whether is different from zero or not at the 0.05 level of significance. State the hypotheses, the value of -test statistic, p-value, and your conclusion.

A slope of zero says y doesn't tend to change linearly when x changes. In other words, there is no linear association the two variables if B1 = 0 

T = b1-0/se(b1)

# Ho: B1 = 0 
# Ha: B1 is not equal zero 

The P-value is the probability of observing a value for the test statistic as contradictory to the null hypothesis and supportive of the alternative hypothesis as the one observed in the sample data.

So, our P value is less than the significance level of .05 so we reject the null hypothesis that the slope is zero.

```{r}

```

# 8. 

```{r}
ci90 = predict(fittedmodelone, data.frame(Hour = c(0,1,2,4,5,6,7,
                                            8,9,10,11,12,13,
                                            14,15,16,17,18,19,20,
                                            21,22,23,24 )), 
        level=0.90,interval="confidence")

plot(sqrt(train$traffic_volume)~train$Hour, main="Regression line with 90% Confidence interval line",xlab="Hour",
        ylab="Sqrt of traffic Volume", )
abline(fittedmodelone,col="red")
lines(ci90[,2],col="blue")
lines(ci90[,3],col="blue")
```

# 9.

There is a curved linear pattern in the residuals
```{r}
plot(fittedmodelone$residuals~fittedmodelone$fitted.values,
     xlab="Predicted",ylab="Residual", main = "Predicted in Relation to residuals")
```

The model is violating the normal distribution assumption  
```{r}
qqnorm(fittedmodelone$residuals)
qqline(fittedmodelone$residuals, col = "red")
```


### Multiple Linear Regression - Part Two

## Model one 

# 1.
```{r}
fittedmodelthree  = lm(sqrt(train$traffic_volume) ~ train$Hour + train$clouds_all+
                      train$temp , data = train)
```

# 2. 

Intercept is of 2.35 is the predicted number of traffic volume when b1 (hours os zero 

B1 is 1.23 positive change in the traffic volume when hour increases one unit 

B2 is .03 positive change in traffic volume while holding the other predictors constant 

B3 is .13 positive change in traffic volume while holding the other predictors constant 

```{r}
summary(fittedmodelthree)
```

# 3.
SST is 22322115 + 28446 + 80230 + 9511832 
     
SSR is 22322115 + 28446 + 80230
     
SSE is 9511832

SSR df is 3 
    
SSE df is 28941

```{r}
anova(fittedmodelthree)
```


# 4. 
H0: Hour, Clouds, and Temp = 0 


Ha: One predictor not equal to 0 


Numerator df = 3 / Denominator = 28941


F statistic  2374 > 2.61 f crit (.05 level)


Reject null that there is no overall linear relationship in the model
```{r}

```

# 5. 


Ho: temp = 0 


Ha: temp not equal to 0 


The p value is less that alpha level of 0.05 


There isn't enough evidence so we reject the null 

*test statistics are shown in the output*

```{r}

```


## Model two 


# 6. 

The categorical variable is getting read in as a factor during preprocessing so we can incorporate it into the model. 
```{r}
fittedmodelfour  = lm(sqrt(train$traffic_volume) ~ train$Hour *train$weather_main,
                      data = train)
```

# 7. 

Predicted volume =  37.20 + 1.25(hour) +/-  bj coef(weather description) +/- bj(hour * weather_description)
```{r}
summary(fittedmodelfour)
```

# 8. 

Categorical term is the variable weather_main. For example in the regression if its predicting a point that has drizzle the equation would like intercept(traffic volume ) + hours(a number ) + clouds(zero) - drizzle(a number) - fog(zero) + haze(zero)...etc 

if its predicting a point that has Fog the equation would like intercept(traffic volume ) + hours(a number ) + clouds(zero) - drizzle(zero) - fog(a number) + haze(zero)...etc 

The intercation term would be hour * weather_main

# 9.

Yes because the R2 went down with having new predictors with high p values assocatied with them. This means that the new coefficents don't add to bettering the model. 

# Model three 

```{r}
fittedmodelfive = lm(sqrt(train$traffic_volume) ~. 
                     - train$Minutes - train$Seconds, data = train )
```

```{r}
summary(fittedmodelfive)
```

# Backward Selection 

```{r}
backselect  = step(fittedmodelfive,direction="backward")
```

# Best model 
sqrt(train$traffic_volume) ~ temp + rain_1h + clouds_all + weather_description + Month + Hour
```{r}
summary(backselect)

```



# Conclusion


With transformations, adding more predictors with hypothesis testing on those predictors, the fitted regression line became better at predicting future values of the hourly data. The r2 went from 12% in a model with no transformations to a r2 value of 21%. I would not recommend using a linear regression on these variables. Splines or Artificial Neural Network would work better in predicting values. An Artificial Neural Network was used into compress the data for anomaly detection.  






















